---
title: "NRC Sentiment_Analysis"
author: "Rob Wells"
date: "2024-8-25"
output: html_document
---

# Jour389/689 Fall 2024

Follow this script and use this dataset instead:
https://github.com/wellsdata/CompText_Jour/raw/main/data/black_press_extracted_text_june_22_2024.csv

load the following libraries: tidyverse, textdata, tidytext, quanteda, rio

```{r}
library(tidyverse)
library(textdata)
library(tidytext)
library(rio)
library(scico)



```


#import "articles_oct_19.csv" as a dataframe

```{r}
#import df created from sequence below
lynch <- read_csv("../data/articles_oct_19.csv")

```

#Tokenize sentence into a df, remove stopwords


```{r}

text_tokenized <- lynch %>%
  select(sentence) %>%
  mutate(sentence = str_replace_all(sentence, "- ", "")) %>%
  unnest_tokens(word, sentence) %>%
  filter(!word %in% stop_words$word) %>%
  filter(word != "stories_corpus") %>%
  filter(!grepl('[0-9]', word))


text_tokenized
```


# Count the words in descending order
```{r}
# Word Count

text_word_ct <- text_tokenized %>%
  count(word, sort=TRUE)

text_word_ct

```

# NRC Sentiment

NRC Lexicon on Whole Corpus
"The nrc lexicon categorizes words in a binary fashion (“yes”/“no”) into categories of positive, negative, anger, anticipation, disgust, fear, joy, sadness, surprise, and trust."
```{r}
# cite this lexicon
#1install.packages("textdata")
nrc_sentiments <- get_sentiments("nrc")
afinn_sentiments <- get_sentiments("afinn")
```

#Count the NRC sentiments
```{r}
text_sentiments <- text_tokenized %>%
  inner_join(nrc_sentiments)

not_sentiments <- text_tokenized %>%
  anti_join(nrc_sentiments)
```

### Join the NRC Sentiments with the tokenized data


```{r}

sentiments_all <- text_tokenized %>%
  inner_join(nrc_sentiments) 

```

### Count Overall Sentiment with NRC

```{r}

sentiments_all <- text_tokenized %>%
  inner_join(nrc_sentiments) %>%
  count(sentiment, sort = TRUE) %>%
  mutate(pct_total = round(n/sum(n), digits=2))

sentiments_all
```

## Use ggplot to chart Sentiments with the tokenized data

```{r}
sentiments_all %>%
  ggplot(aes(x = sentiment, y = n, fill=n)) +
  geom_col() +
  labs(title = "Sentiments in Lynching Data",
       caption = "Graphic by Lang")


nrc_plot <- sentiments_all %>%
  ggplot(aes(x = sentiment, y = n, fill=n)) +
  geom_col(position="dodge") +
  labs(title = "Total Sentiment in Lynching Coverage",
       subtitle="",
       y="Score",
       x="Word")

nrc_plot + scico::scale_fill_scico(palette = "vik")

nrc_plot
```



# Create a new dataframe just with the NRC "anger" sentiment
```{r}
nrc_anger <- nrc_sentiments %>%
  filter(word = "anger")

nrc_anger

```

